Leveraging open platforms to manage and distribute streams of versioned data

### 2017-07-11 ms.tex
more editing

@@

....

(old content unchanged)

One of the key problems in switching from a static to a dynamic data
world, is that it's currently hard to set up and maintain a system like
this. Ideally using a versioned database should be easy for both the
user and the data maintainer. This is especially crucial is the data
maintainers are busy. Moreover, old versions need to be preserved in a
stable and accessible way.

To try and achieve to above goals, we built a system to distribute
versions of our datasets. We think that with this infrastructure, it is
relatively easy to set up and maintain a versioned system with any
small-to-medium sized dataset.

In our system, the data is downloaded only once per computer, available
for offline access, easy to port to other platforms, and can get a doi
when desired (i.e.~when it's used for a publication). All old versions
are accessible.

\subsection{How much expertise is needed to setup and use?}

To use a versioned dataset: Very very basic \texttt{R}. With 4 lines of \texttt{R} code, you can
have the data downloaded and loaded into R and ready for use.

To set a system like this up: Basically there are two general skills required: 1) understanding
version control via \texttt{git} and 2) writing a very simple R package.
From the prolific Hadley Wickham there is a
\href{http://r-pkgs.had.co.nz/git.html}{very clear introduction to git}
and \href{http://r-pkgs.had.co.nz/intro.html}{an instruction to the
basics of R packages}.

With those two skills the rest should be relatively easy. To set up a
living dataset you need to provide:
\begin{enumerate}
  \item a github repository name (e.g.~`wcornwell/taxonlookup`, `traitecoevo/baad.data`).
  \item A single file that will be the versioned dataset (e.g. \texttt{plant\_lookup.csv} or \texttt{baad\_data.zip}).
  \item A function to load that data into R:(e.g.~a simple wrapper around \texttt{read.csv}).
\end{enumerate}

>>>>

[MWP: moved this from discussion]
To set up a
living dataset you need to provide:
\begin{enumerate}
  \item a github repository name (e.g.~`wcornwell/taxonlookup`, `traitecoevo/baad.data`).
  \item A single file that will be the versioned dataset (e.g. \texttt{plant\_lookup.csv} or \texttt{baad\_data.zip}).
  \item A function to load that data into R:(e.g.~a simple wrapper around \texttt{read.csv}).
\end{enumerate}

@@

%(old content unchanged)
%
%One of the key problems in switching from a static to a dynamic data
%world, is that it's currently hard to set up and maintain a system like
%this. Ideally using a versioned database should be easy for both the
%user and the data maintainer. This is especially crucial is the data
%maintainers are busy. Moreover, old versions need to be preserved in a
%stable and accessible way.
%
%To try and achieve to above goals, we built a system to distribute
%versions of our datasets. We think that with this infrastructure, it is
%relatively easy to set up and maintain a versioned system with any
%small-to-medium sized dataset.
%
%In our system, the data is downloaded only once per computer, available
%for offline access, easy to port to other platforms, and can get a doi
%when desired (i.e.~when it's used for a publication). All old versions
%are accessible.
%
%
%
%\subsection{How much expertise is needed to setup and use?}
%
%To use a versioned dataset: Very very basic \texttt{R}. With 4 lines of \texttt{R} code, you can
%have the data downloaded and loaded into R and ready for use.
%
%To set a system like this up: Basically there are two general skills required: 1) understanding
%version control via \texttt{git} and 2) writing a very simple R package.
%From the prolific Hadley Wickham there is a
%\href{http://r-pkgs.had.co.nz/git.html}{very clear introduction to git}
%and \href{http://r-pkgs.had.co.nz/intro.html}{an instruction to the
%basics of R packages}.
